---
title: "P8105 Homework 6"
author: "sarah_8105"
output: github_document
---


This is my sixth homework assignment for P8105.

```{r libraries}
library(tidyverse)
library(rvest)
library(readr)
library(httr)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

In this first code chunk, I pull the Washington Post data on homicides in 50 large US cities. I create a `city_state` variable that concatenates the city and state that the homicide occurred in and remove an observation that was erroneously attributed to Tulsa, AL. 

```{r p1_import, cache = TRUE}
homicide_df = read_csv(file = "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv") %>%
  mutate(
    city_state = str_c(city, state, sep = "_")
  ) %>%
  select(city_state, disposition) %>%
  filter(city_state != "Tulsa_AL")
```




## Problem 2

Problem 2 concerns data from a longitudinal study that includes control and experimental study arms. In the following code chunk, I import data on each study participant that have been stored as separate csv files. I create an `import_csv` function that is used to import and clean each csv file and I iterate over all csv file using `purrr::map`. I also create variables representing the study arm and study ID of each participant. Since study ID is duplicated over the study arms and is therefore not unique, I also create a `unique_id` variable that can be used to identify each participant.

```{r p2_import, message = FALSE}
p2_df = 
  tibble(
    file_name = list.files(path = "./data/")
    )

import_csv = function(file) {
  read_csv(file) %>%
  janitor::clean_names() %>%
    pivot_longer(
      week_1:week_8,
      names_to = "week",
      names_prefix = "week_",
      values_to = "observation"
    )
}

p2_df = p2_df %>%
  mutate(
    unique_id = seq.int(nrow(p2_df)),
    study_id = str_sub(file_name, 5, 6),
    study_arm = ifelse(str_detect(file_name, "con"), "Control", "Experimental"),
    data = map(str_c('./data/',file_name), import_csv),
    ) %>%
  select(-file_name) %>%
  unnest()
```



