---
title: "P8105 Homework 6"
author: "sarah_8105"
output: github_document
---


This is my sixth homework assignment for P8105.

```{r libraries}
library(tidyverse)
library(rvest)
library(readr)
library(httr)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## Problem 1

In this first code chunk, I pull the Washington Post data on homicides in 50 large US cities. I create a `city_state` variable that concatenates the city and state that the homicide occurred in and remove an observation that was erroneously attributed to Tulsa, AL. 

```{r p1_import, cache = TRUE}
homicide_df = read_csv(file = "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv") %>%
  mutate(
    city_state = str_c(city, state, sep = "_")
  ) %>%
  select(city_state, disposition) %>%
  filter(city_state != "Tulsa_AL")
```




## Problem 2

Problem 2 focuses on understanding the effects of different variables on low birth weight. The data set consists of ~4000 children and factors related to their birth and their parents. In this first code chunk, I import and explore the data to assess distributions and missingness.


```{r bwt_import}
bwt_df = 
  read_csv("Data/birthweight.csv") %>%
  janitor::clean_names() %>%
  mutate(
    babysex = as_factor(babysex),
    frace = as_factor(frace),
    malform = as_factor(malform),
    mrace = as_factor(mrace),
    smoking_status = as_factor(if_else(smoken == 0, 0, 1)),
    menarche = if_else(menarche == 0, NA_real_, menarche)
  )  %>%
  select(-pnumlbw, -pnumsga)

summary(bwt_df)

```

There appears to be no missing data in this data set, but it's possible that some of the zeros represent missing data. For example, an age of menarchy of 0 is likely missing or erroneous and was therefore set to missing. However, some zeros are likely to be valid entries. For example, an average number of cigarettes smoked of 0 likely indicates that the mother was not a smoker. This variable was converted into a categorical smoking status variable of smoker vs. non-smoker. The number of prior low birth weight or small for gestational age babies is 0 for all subjects, so these variables will be uninformative for our models and were dropped from the data set. 

In the next code chunk I compare three models:
* Model 1: main effects of length at birth and gestational age only
* Model 2: including predictors of head circumference, length, sex, and all interactions between these variables
* Model 3: predictors in model 1 & 2, in addition to factors that have been demonstrated as significant predictors of low birth weight in the literature, including mother's race, baby's sex, smoking status, weight gain, and mother's pre-pregnancy BMI

```{r bwt_models}
model1 = lm(bwt ~ blength + gaweeks, data = bwt_df)
broom::glance(model1)
broom::tidy(model1)

bwt_df %>%
  modelr::add_predictions(model1) %>%
  modelr::add_residuals(model1) %>%
  ggplot(aes(x = pred, y = resid)) +
    geom_point(alpha = .5)

model2 = lm(bwt ~ bhead + blength + babysex + babysex*bhead*blength, data = bwt_df)
broom::glance(model2)
broom::tidy(model2)

bwt_df %>%
  modelr::add_predictions(model2) %>%
  modelr::add_residuals(model2) %>%
  ggplot(aes(x = pred, y = resid)) +
    geom_point(alpha = .5)
  
model3 = lm(bwt ~ blength + gaweeks + bhead + blength + babysex + babysex*bhead*blength + mrace + babysex + smoken + momage + ppbmi + wtgain, data = bwt_df)
broom::glance(model3)
broom::tidy(model3)

bwt_df %>%
  modelr::add_predictions(model3) %>%
  modelr::add_residuals(model3) %>%
  ggplot(aes(x = pred, y = resid)) +
    geom_point(alpha = .5)

```


## Problem 3

In this problem, I use 2017 Central Park weather data. First, I import the data using the P8105 code.

```{r weather_import, cache = TRUE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Next, I run a simple linear regression of tmax on tmin using 5000 bootstrap samples. I also calculate log(β^0 ∗ β^1) for each regression output, plot the distribution of this estimate for all 5000 samples, and calculate the mean and 95% confidence interval. 

```{r weather_boot}
boot_strap_results = 
  weather_df %>%
  modelr::bootstrap(n = 5000) %>%
  mutate(
    models = map(strap, ~lm(tmax ~ tmin, data = .x) ),
    results = map(models, broom::tidy),
    r2 = map(models, broom::glance)) %>%
  select(-strap, -models) 

terms = 
  boot_strap_results %>%
  unnest(results) %>%
  select(.id, term, estimate) %>%
  pivot_wider(
    names_from = "term",
    values_from = "estimate"
  ) %>%
  janitor::clean_names() %>%
  mutate(
    estimate = log(intercept * tmin)
  ) 


terms %>%
  ggplot(aes(x = estimate)) + 
  geom_density() +
  labs(
    title = "Distribution in log(β^0 ∗ β^1) Using 5000 Bootstrap Samples",
    x = "log(β^0 ∗ β^1)",
    y = "Density"
  )

terms %>%
  summarize(
    Mean = mean(estimate),
    LCL = quantile(estimate, 0.025),
    UCL = quantile(estimate, 0.975)
    ) %>% 
  knitr::kable(digits = 2)
```

In this last code chunk, I plot the distribution of the R^2 for all 5000 samples and calculate the mean and 95% confidence interval.  

```{r p3_plots}
r2 = 
  boot_strap_results %>%
  unnest(r2) 


r2 %>%
  ggplot(aes(x = r.squared)) +
  geom_density() +
  labs(
    title = "Distribution in R^2 Using 5000 Bootstrap Samples",
    x = "R^2",
    y = "Density"
  )

r2 %>%
  summarize(
    Mean = mean(r.squared),
    LCL = quantile(r.squared, 0.025),
    UCL = quantile(r.squared, 0.975)
    ) %>% 
  knitr::kable(digits = 3)
```
